---
title: "PS270 Final Project"
author: "Landon Johnson"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to lobbying and how much certain fields/companies spend in lobbying. I believe it would be fascinating to see who/what companies have political influence. By analyzing this data, we will also be able to see what politicians support what companies/industries.




## Project Paragraph
My research question is: How does the amount of money spent by companies on lobbying influence the concentration of lobbying expenditures within specific industries? I hypothesize that industries with high regulatory stakes such as health, finance, and energy will have a greater concentration of lobbying expenditures among a few top companies than less regulated ones. This hypothesis is rooted in the theory of political influence, suggesting that companies within highly regulated industries have a strong incentive to invest disproportionately in lobbying to secure favorable policy outcomes. The institutional structure of these industries often creates a high-stakes environment where policy changes can significantly impact profitability, driving a concentration of lobbying efforts within top firms. The independent variable is the amount of money spent on lobbying by companies, which can be measured in dollars. I have data that provides lobbying expenditures by industry and company over time, allowing me to observe variations in spending across different companies and sectors. The dependent variable is the concentration of lobbying expenditures within each industry, measured as the proportion of total lobbying spending accounted for by the top companies in that industry. This variable can be calculated by identifying the percentage of total lobbying dollars spent by the top companies in each industry. If the data shows that a small number of companies within highly regulated industries account for a large portion of total lobbying expenditures, this would support my hypothesis that these industries have a more concentrated lobbying effort driven by high stakes. Conversely, if lobbying expenditures are spread more evenly across companies within highly regulated industries, or if non-regulated industries show similar concentration patterns, this would counter my hypothesis, suggesting that regulatory stakes do not drive concentrated lobbying.


## Project Data Visualization 

```{r}
library(dplyr)
library(ggplot2)
library(knitr)
```



```{r}
getwd()
```

```{r}
top_spenders <- read.csv("Top Spenders.csv")
ranked_sectors <- read.csv("Ranked Sectors.csv")
top_issues <- read.csv("Top Issues.csv")
```

```{r}
head(top_spenders)
head(ranked_sectors)
head(top_issues)
```


```{r}
colnames(top_spenders)
```





```{r}
colnames(ranked_sectors)
colnames(top_spenders)
```


```{r}
ranked_sectors$Total <- as.numeric(gsub("[$,]", "", ranked_sectors$Total))
```

```{r}
ggplot(data = ranked_sectors, aes(x = reorder(Sector, -Total), y = Total)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black", width = 0.7) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M"),  
                     expand = expansion(mult = c(0, 0.1))) +  
  labs(
    title = "Distribution of Total Lobbying Expenditures by Sector",
    x = "Sector",
    y = "Total Expenditure ($ Millions)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", margin = margin(b = 15)),
    panel.grid.major.y = element_line(color = "grey80"), 
    panel.grid.minor.y = element_blank() 
  )
```

## Data Visualization and Analysis




```{r}

ranked_sectors$Total_Millions <- ranked_sectors$Total / 1e6

ranked_sectors[order(-ranked_sectors$Total_Millions), c("Sector", "Total_Millions")]

```

         
      
```{r}
library(knitr)

final_table <- data.frame(
  Sector = c(
    "Communic/Electronics",
    "Energy/Nat Resource",
    "Finance/Insur/RealEst",
    "Health",
    "Misc Business"),
  Top_Spender = c(
    "NCTA - The Internet & Television Assn",
    "American Petroleum Institute",
    "National Assn of Realtors",
    "Pharmaceutical Research & Manufacturers",
    "U.S. Chamber of Commerce"),
  Total_Sector_Expenditure_M = c(439.46, 315.73, 473.32, 562.33, 430.49),
  Top_Spender_Contribution = c(5.46, 6.61, 11.28, 11.29, 4.95))

kable(final_table, format = "html", col.names = c(
  "Sector", "Top Spender", "Total Sector Expenditure ($M)", "Top Spender Contribution (%)"
), align = "c")

```
The chart and table collectively provide important insights into how lobbying expenditures influence legislative outcomes, helping to address the research question: How does the amount of money companies spend on lobbying affect the voting behavior of politicians on legislative decisions that impact those industries? The bar chart reveals the concentration of lobbying expenditures by sector, highlighting that the Health and Finance/Insurance/Real Estate sectors dominate, with totals exceeding $500 million each. These industries, being highly regulated, are likely focusing their lobbying efforts on shaping policies directly affecting their operations. Complementing the chart, the table identifies the top spenders within these sectors, such as Pharmaceutical Research & Manufacturers, which contributes 11.29% of lobbying expenditures in the Health sector, and the National Assn of Realtors, responsible for 11.28% in Finance/Insurance/Real Estate. These data points underscore the dominance of a few organizations within their respective sectors, suggesting that companies investing heavily in lobbying are more likely to influence legislative decisions. While the chart and table do not directly measure voting behavior, they imply that industries spending the most on lobbying gain greater access to policymakers and likely influence legislative priorities. The findings suggest that financial investment in lobbying correlates with significant political influence, aligning with the hypothesis that industries with higher lobbying expenditures are more successful in shaping legislative outcomes in their favor.



